# Thordis Ingadottir's "Project 2 in IODS, data wrangling and regression"

*Describe the work you have done this week and summarize your learning.*

- Describe your work and results clearly. 
- Assume the reader has an introductory course level understanding of writing and reading R code as well as statistical methods.
- Assume the reader has no previous knowledge of your data or the more advanced methods you are using.

```{r}
date()
```

## This week:
I have been familiarising myself with R by doing the exercises in Datacamp. This included learning how to load data from the web, writing and reading locally, plotting, linear regression functions etc. This weeks project follows here:

1. Load the data locally 
```{r}
learning2014 <- read.csv("./data/learning2014.csv", header = T, sep = ",")
learning2014
```
As can be seen this loaded a dataframe with student data. The headers list 'variables':
```{r}
head(learning2014)
```
with data about the students, their grades (points) and averaged answers to different categories of questions (the variables deep, stra, and surf).
The dimensions of the dataframe are 7 such variables, times 166 "observations", i.e. 166 students.
```{r}
str(learning2014)
dim(learning2014)
```

