---
output:
  html_document: default
  pdf_document: default
---
# Thordis Ingadottir's "Project 2 in IODS, data wrangling and regression"

*Describe the work you have done this week and summarize your learning.*

- Describe your work and results clearly. 
- Assume the reader has an introductory course level understanding of writing and reading R code as well as statistical methods.
- Assume the reader has no previous knowledge of your data or the more advanced methods you are using.

```{r}
date()
```

## This week:
I have been familiarising myself with R by doing the exercises in Datacamp. This included learning how to load data from the web, writing and reading locally, plotting, linear regression functions etc. This weeks project follows here:

## 1. Load the data locally 
```{r}
learning2014 <- read.csv("./data/learning2014.csv", header = T, sep = ",")
learning2014
```

As can be seen this loaded a dataframe with student data. The headers list 'variables':
```{r}
head(learning2014)
```
with data about the students, their grades (points) and averaged answers to different categories of questions (the variables deep, stra, and surf).
The dimensions of the dataframe are 7 such variables, times 166 "observations", i.e. 166 students.
```{r}
str(learning2014)
dim(learning2014)
```

## 2. Graphical overview of the data:
Different ways of presenting quick overviews of the data include 'pairs' and 'ggpairs':
```{r}
library(GGally)
library(ggplot2)
# pairs(learning2014[-1])
p <- ggpairs(learning2014, mapping = aes(col = gender, alpha = 0.3), lower = list(combo = wrap("facethist", bins = 20)))
p
```

The pairs function works on my console, but in the markup file it complains about non-numeric argument being supplied to 'pairs'! Exactly the same function call. Apparently my console is more forgiving

There's not an awful lot I can say about this data. In terms of correlation with points, the age is very clustered at the low end, attitude seems to have a decent correlation with points (surprise?). I can't say there seems to be much of a correlation in the question variables deep, stra, and surf, except perhaps a little for stra. Maybe we'll find out in the next section.

Summary of the variables (except for gender, that isn't informative as the variable is of string type) follows:
```{r}
summary(learning2014$age)
summary(learning2014$attitude)
summary(learning2014$deep)
summary(learning2014$stra)
summary(learning2014$surf)
summary(learning2014$points)
```
The output is self explanatory, min and max values, median and mean, and the 1st Qu. and 3rd Qu. give a very rough idea of the width of the distribution (how sharp the presumed peak at the mean).

## 3. Choose three explanatory variables and fit a regression model with points as the target variable.
I will choose attitude, stra and surf as explanatory variables (in the Datacamp these had the highest correlations with points), with points as the target variable. Create a regression model m:
```{r}
m <- lm(points ~ attitude + stra + surf, data = learning2014)
summary(m)
```
I will refrain from too much interpretation of these model parameters. The attitude does show a good fit, with probability of 'null hypothesis' in the last column being very low. Also a low std. error. The other two variable are less impressive fits with the points, although surf seems to be the worst without doubt.

## 4. Explain the relationship between the chosen explanatory variables and the target variable (interpret the model parameters).
This was addressed above. Attitude has the best fit. It is interesting that removing the surf variable improves the fit for stra, as shown below:
```{r}
m2 <- lm(points ~ attitude + stra, data = learning2014)
summary(m2)
```
## 5. Diagnostic plots.
These are the diagnostic plots. They look rather similar for the three and two explanatory variables.
```{r}
par(mfrow = c(2,2))
plot(m, which = c(1,2,5))
```

The residuals look reasonably random. The QQ-plot does indicate problems at both ends and the leverage plot is a little sparse to the right hand side. The model assumptions are that errors are normally distributed and independent of explanatory variable.

This exercise has been very educational, as my prior statistics knowledge is not very extensive... and R seems very powerful!

